---
title: "Problem Set 2 "
author: "Yunqiu(Julie) Li"
date: "Spring 2019"
output:
  word_document: default
  html_document:
    df_print: paged
---


```{r, message = FALSE, warning= FALSE}
library(forecast)
library(ggplot2)
```

# Problem 1: For this problem load the Logan airport data set from the lecture csv file.
## Create time series for the domestic and international passengers.
```{r}
logan.data <- read.csv("LoganPassengers.csv")
logan.data <- logan.data[logan.data$total<5000000,]
logan.dom <- ts(logan.data$domestic, start = c(2002,10), freq = 12)
logan.int<- ts(logan.data$international, start = c(2002,10), freq = 12)
```

## Plot both of these lines in the same plot. (Hint, use the plot command, and then the lines command for this. Type help(lines) for help on the lines command.
```{r}
plot(logan.dom,xlab="Year",ylab="Passengers",bty="l", ylim=range(90000, 1500000), col = "blue")
lines(logan.int, col = "red")
legend("topleft",legend=c("Domestic", "International"),col=c("blue","red"),lty=1:1, cex = 0.7)
grid()
```

## Describe any similarities and differences between the two.
Similarities:

* The patterns for domestic passengers and international passenger have an upward trend.

* The number of domstic passengers and international passengers both have seasonality variations.

* The seasonality effects tend to increase over time.

Differences:

* The overall scale of internaiontal passengers is way larger than the overall scale of domestic passengers.

* International passengers experience a larger seasonal violatility.

* There is a significant dip in the number of international passengers at the beginning of 2019.

* Within a year, the patterns of fluctuation in domestic passengers and international passengers are different, which indicates distinct customer travel preference.


## Repeat this plot for a short window from 2014 to the end of the data set.
```{r}
logan.dom.zoom <- window(logan.dom,start=c(2014,1))
logan.int.zoom <- window(logan.int,start=c(2014,1))
plot(logan.dom.zoom,xlab="Year",ylab="Passengers",bty="l", ylim=range(90000, 1600000), col = "blue")
lines(logan.int.zoom, col = "red")
legend("topleft",legend=c("Domestic", "International"),col=c("blue","red"),lty=1:1, cex = 0.7)
grid()
```

# Problem 2: For this problem use the global temperature data that we used in class. Load and convert to an R ts object as we did in class. Define the last 20 years of data as the validation set, and the remaining data the training set.
## Plot the full data set.
```{r}
temp.data <- read.csv("globalTemps.csv")
temp.ts <- ts(temp.data$temp,start=1880,freq=1)
temp.smooth <- ts(temp.data$Lowess5,start=1880,freq=1)
stepsAhead <- 20
nTrain <- length(temp.ts) - stepsAhead
temp.train <- window(temp.ts, start = 1880, end = 1880-1+nTrain)
temp.test <- window(temp.ts, start = 1880+nTrain, end = 1880-1+nTrain + stepsAhead)
plot(temp.ts,xlab="Year",ylab="Global surface temp")
lines(temp.smooth, col = "red")
grid()
```

## Now estimate a model with just a linear trend in it using the training data. Report the mean squared forecast error and mean absolute error for both the training and validation samples.
```{r}
temp.lm <-  tslm(temp.train ~ trend)
temp.lm.pred <- forecast(temp.lm, h = stepsAhead, level = 0)
# validation forecast errors 
valid.err <- temp.test - temp.lm.pred$mean
# train forecast errors 
train.err <- temp.train - temp.lm$fitted.values

# train mean squared forecast error
mse.train <- mean( (train.err)^2 )
# validation mean squared forecast error
mse.test <- mean( (valid.err)^2 )
# train mean absolute error
mae.train <- mean(abs(train.err))
# validation mean absolute error
mae.test <- mean(abs(valid.err))

sprintf("The training sample MSE is %s", round(mse.train,5))
sprintf("The validation sample MSE is %s", round(mse.test,5))
sprintf("The training sample MAE is %s", round(mae.train,5))
sprintf("The validation sample MAE is %s", round(mae.test,5))
```

## Plot both parts of the data set along with the fitted model in the training and validation sets.
```{r}
plot(temp.lm.pred,  ylab = "Global surface temp", xlab = "Year", bty = "l",  flty = 2, main = "", ylim=c(-0.5,1))
# plot fitted values in the training period
lines(temp.lm$fitted, lwd = 2)
# plot data in the validation period
lines(temp.test)
lines(temp.test, col="red")
lines(c(1998,1998), c(-0.5, 1),lwd=3,col="red") 
text(1940, 1, "Training",cex=1.25)
text(2010, 1, "Validation",cex=1.25)
grid()
```

## Now change the model to a quadratic trend model. Report the mean squared errors, and the mean absolute error in the training and validation data sets.
```{r}
temp.qm <-  tslm(temp.train ~ trend + I(trend^2))
temp.qm.pred <- forecast(temp.qm, h = stepsAhead, level = 0)

# train forecast errors 
temp.train.err <- temp.train - temp.qm$fitted.values
# validation forecast errors 
temp.valid.err <- temp.test - temp.qm.pred$mean

# train mean squared forecast error
temp.mse.train <- mean( (temp.train.err)^2 )
# validation mean squared forecast error
temp.mse.test <- mean( (temp.valid.err)^2 )
# train mean absolute error
temp.mae.train <- mean(abs(temp.train.err))
# validation mean absolute error
temp.mae.test <- mean(abs(temp.valid.err))

sprintf("The training sample MSE is %s", round(temp.mse.train,5))
sprintf("The validation sample MSE is %s", round(temp.mse.test,5))
sprintf("The training sample MAE is %s", round(temp.mae.train,5))
sprintf("The validation sample MAE is %s", round(temp.mae.test,5))
```

# Problem 3: For this problem load the US unemployment data. Write R code to load the data set, and convert it to a ts object. Let your training set go from Jan 1985 to Dec 2009. The remaining data after Dec 2009 is the validation set. Fit a quadratic trend model to the training data, and use it to forecast the validation set.
```{r}
unemploy.data <- read.csv("UNRATE.csv")
unemploy.ts <- ts(unemploy.data$UNRATE,start=c(1948,1),freq=12)
unemploy.train <- window(unemploy.ts, start = c(1985,1), end = c(2009,12))
unemploy.valid <- window(unemploy.ts, start = c(2010,1))
unemploy.qm <-  tslm(unemploy.train ~ trend + I(trend^2))
unemploy.qm.pred <- forecast(unemploy.qm, h = 103, level = 95)
```

## Plot the quadratic trend forecast in the training and validation period, and also show the 95 percent confidence bands for the forecast.
```{r}
plot(unemploy.qm.pred, ylab = "Unemployment Rate", bty='l', xlab = "Year", main = "", flty = 2)
lines(unemploy.qm$fitted, lwd = 2)
lines(unemploy.valid, col="red")
lines(c(2010.1, 2010.1), c(0,25), lwd=3, col="red") 
text(1998, 12, "Training", cex=1.25)
text(2015, 12, "Validation", cex=1.25)
grid()
```

## Estimate the mean squared error and absolute forecast error for both the training and validation periods.
```{r}
unemploy.train.err <- unemploy.train - unemploy.qm$fitted.values
unemploy.valid.err <- unemploy.valid - unemploy.qm.pred$mean
unemploy.mse.train <- mean(unemploy.train.err^2)
unemploy.mse.test <- mean(unemploy.valid.err^2)
unemploy.mae.train <- mean(abs(unemploy.train.err))
unemploy.mae.test <- mean(abs(unemploy.valid.err))

sprintf("The training sample MSE is %s", round(unemploy.mse.train,5))
sprintf("The validation sample MSE is %s", round(unemploy.mse.test,5))
sprintf("The training sample MAE is %s", round(unemploy.mae.train,5))
sprintf("The validation sample MAE is %s", round(unemploy.mae.test,5))
```

## Repeat this with a naive forecast starting at the end of the training period. Use the naive() function, and remember to plot the 95 percent confidence bands.
```{r}
unemploy.qm.naive <- naive(unemploy.train, h = 103, level = 95) 
plot(unemploy.qm.naive, ylab = "Unemployment Rate", bty='l',xlab = "Year",main = "", flty = 2)
lines(unemploy.valid, col="red")
lines(c(2010.1, 2010.1), c(0,12),lwd=3,col="red") 
text(1998, 12, "Training",cex=1.25)
text(2015, 12, "Validation",cex=1.25)
grid()

naive.unemploy.train.err <- unemploy.train - unemploy.qm.naive$fitted
naive.unemploy.valid.err <- unemploy.valid- unemploy.qm.naive$mean
unemploy.naive.mse.train <- mean(naive.unemploy.train.err^2, na.rm=TRUE)
unemploy.naive.mse.test <- mean(naive.unemploy.valid.err^2)
unemploy.naive.mae.train <- mean(abs(naive.unemploy.train.err), na.rm=TRUE)
unemploy.naive.mae.test <- mean(abs(naive.unemploy.valid.err))
sprintf("The training sample naive MSE is %s", round(unemploy.naive.mse.train,5))
sprintf("The validation sample naive MSE is %s", round(unemploy.naive.mse.test,5))
sprintf("The training sample naive MAE is %s", round(unemploy.naive.mae.train,5))
sprintf("The validation sample naive MAE is %s", round(unemploy.naive.mae.test,5))
```
